Stage-structured evolutionary demography for simultaneous hermaphrodites with partial selfing

### 2020-07-03 notes/notes-SA-partSelf-wDemography.tex
testing simulation with new Atilde structure
updates to notes with new Atilde

\noindent Individuals are jointly classified by stage $(1, \ldots , \omega)$ and genotype $(1, \ldots , g)$. Each genotype is characterized by a matrix of transition probabilities (including survival) and a matrices of reproductive output through male and female function. These matrices can include time variation or nonlinearities reflecting the environment or density dependence, although thes complexities will not be addressed here. Each stage contributes offspring to genotypes at the next time step according to matrices that are determined by the mating system and the population structure.

@@

\begin{equation}
	\mathcal{N} \sim \left[
						\begin{array}{ccc}
							n_{1,1}      &  \cdots & n_{1,g} \\
							\vdots       &         & \vdots  \\
							n_{\omega,1} &  \cdots & n_{\omega,g} \\
						\end{array} \right],
\end{equation}

@@

\begin{equation}
	\tilde{\mbf{n}}(t) =  \left[
								\begin{array}{c}
									\mbf{n}_{AA}(t) \\
									\mbf{n}_{Aa}(t) \\
									\mbf{n}_{aa}(t) \\
						\end{array} \right].
\end{equation}

@@

\begin{equation}
	\tilde{\mbf{A}}[\tilde{\mbf{n}}] = 
			\underbrace{\left(
			\begin{array}{c|c}
				\mcal{U}^S & \mbf{0} \\ \hline
				\mbf{0} & \mcal{U}^X \\
			\end{array} \right)}_{\tilde{\mbf{U}}} + 
			\underbrace{\left(
			\begin{array}{c|c}
				\mcal{F}^S(\tilde{\mbf{p}}) (1 - \delta) & \mbf{0} \\ \hline
				\mbf{0} & \mcal{F}^X(\tilde{\mbf{p}}) \\
			\end{array} \right)}_{\tilde{\mbf{F}}}
\end{equation}

@@

\noindent and corresponding matices $\mbb{U}^X$, $\mbb{F}^S$, $\mbb{F}^X$. Because individuals cannot change their genotype once they are born, the survival matrices are block diagonal; $\mcal{U}^S = \mbb{U}^S$ and $\mcal{U}^X = \mbb{U}^X$. We also construct

@@

\begin{align*}
	\tilde{\mbf{A}}[\tilde{\mbf{n}}] = 
		& \underbrace{\left(
			\begin{array}{ccc|ccc}
				\mcal{U}^S_{AA} & \mbf{0} & \mbf{0} & \mbf{0} & \mbf{0} & \mbf{0} \\ 
				\mbf{0} & \mcal{U}^S_{Aa} & \mbf{0} & \mbf{0} & \mbf{0} & \mbf{0} \\
				\mbf{0} & \mbf{0} & \mcal{U}^S_{Aa} & \mbf{0} & \mbf{0} & \mbf{0} \\ \hline
				\mbf{0} & \mbf{0} & \mbf{0} & \mcal{U}^X_{AA} & \mbf{0} & \mbf{0} \\
				\mbf{0} & \mbf{0} & \mbf{0} & \mbf{0} & \mcal{U}^X_{aa} & \mbf{0}\\
				\mbf{0} & \mbf{0} & \mbf{0} & \mbf{0} & \mbf{0} & \mcal{U}^X_{aa}
			\end{array} \right)}_{\tilde{\mbf{U}}} \\ 
			&+ \underbrace{\left(
			\begin{array}{ccc|ccc}
				C_{AA} (1 - \delta) \mbf{F}_{AA} & \frac{1}{4} C_{Aa} (1 - \delta) \mbf{F}_{Aa} & 0 & \mbf{0} & \mbf{0} & \mbf{0} \\ 
				0 & \frac{1}{2} C_{Aa} (1 - \delta) \mbf{F}_{Aa} & 0 & \mbf{0} & \mbf{0} & \mbf{0} \\
				0 & \frac{1}{4} C_{Aa} (1 - \delta) \mbf{F}_{Aa} & C_{aa} (1 - \delta) \mbf{F}_{aa} & \mbf{0} & \mbf{0} & \mbf{0} \\ \hline
				q^{\prime}_{A} \mbf{F}_{AA} (1 - C_{AA})& \frac{1}{2} q^{\prime}_{A} \mbf{F}_{Aa} (1 - C_{Aa}) & 0 & \mbf{0} & \mbf{0} & \mbf{0} \\
				q^{\prime}_{a} \mbf{F}_{AA} (1 - C_{AA}) & \frac{1}{2} \mbf{F}_{Aa} (1 - C_{Aa}) & q^{\prime}_{A} \mbf{F}_{aa} (1 - C_{aa}) & \mbf{0} & \mbf{0} & \mbf{0}\\
				0 & \frac{1}{2} q^{\prime}_{a} \mbf{F}_{Aa} (1 - C_{Aa})& q^{\prime}_{a} \mbf{F}_{aa} (1 - C_{aa}) & \mbf{0} & \mbf{0} & \mbf{0}
			\end{array} \right)}_{\tilde{\mbf{F}}} \numberthis
\end{align*}

@@

As a first pass at implementing the model, we construct and analyze a genotype $\times$stage-classified model for a hypothetical species with intralocus sexual conflict via the two sex functions. Our hypothetical species has two life stages: juveniles and adults. Suppose that allele $A$ is beneficial for females but detrimental for males, and that allele $a$ has the reverse effect. We also assume that the effects act only during the adult stage through reproductive success. The allele does not affect survival and transition rates, such that

\begin{equation}
	\mbf{U} = \left(
					\begin{array}{cc}
						\sigma_J(1 - \gamma) & 0 \\
						\sigma_J \gamma      & \sigma_A
					\end{array}
				\right),
\end{equation}

@@

\begin{equation}
	\mbf{F}_{ii} = \left(
					\begin{array}{cc}
						0 & f_{ii} \\
						0 & 0
					\end{array}
				\right),
\end{equation}
\noindent and
\begin{equation}
	\mbf{F}^{\prime}_{ii} = \left(
					\begin{array}{cc}
						0 & f^{\prime}_{ii} \\
						0 & 0
					\end{array}
				\right),
\end{equation}

@@

\begin{equation}
	\tilde{\mbf{A}}[\tilde{\mbf{n}}] = 
			\underbrace{\left(
			\begin{array}{c|c}
				\mcal{U}^S & \mbf{0} \\ \hline
				\mbf{0} & \mcal{U}^X \\
			\end{array} \right)}_{\tilde{\mbf{U}}} + 
			\underbrace{\left(
			\begin{array}{c|c}
				\mcal{F}^S(\tilde{\mbf{p}}) C(1 - \delta) & \mbf{0} \\ \hline
				\mbf{0} & \mcal{F}^X(\tilde{\mbf{p}}) \\
			\end{array} \right)}_{\tilde{\mbf{F}}}
\end{equation}

\begin{table}[htbp]
 \centering
 \caption{\bf Fitness expressions for Sexually Antagonistic selection in hermaphrodites}
\begin{tabular}{lccc}
 \toprule
					&  \multicolumn{3}{c}{{\textit{Genotype}}} \\ 
\cline{2-4}
					& $AA$		& $Aa$ 			& $aa$ 		\\ \hline
Female function:	& $f$		& $f - h_f s_f$	& $f - s_f$ \\	
Male function:		& $f - s_m$	& $f - h_m s_m$	& $f$ 		\\	
\hline
\end{tabular}
\end{table}

>>>>

\noindent Individuals are jointly classified by whether they were produced by self-fertilization or outcrossing (denoted by $S$ and $X$), stage $(1, \ldots , \omega)$ and genotype $(1, \ldots , g)$. Each genotype is characterized by a matrix of transition probabilities (including survival) and a matrices of reproductive output through male and female function. These matrices can include time variation or nonlinearities reflecting the environment or density dependence, although thes complexities will not be addressed here. Each stage contributes offspring to genotypes at the next time step according to matrices that are determined by the mating system and the population structure.

@@

\begin{equation}
	\mathcal{N} \sim \left[
						\begin{array}{ccc|ccc}
							n^S_{1,1}      &  \cdots & n^S_{1,g}       &  n^X_{1,1}      &  \cdots & n^X_{1,g} \\
							\vdots       &         & \vdots        &  \vdots       &         & \vdots  \\
							n^S_{\omega,1} &  \cdots & n^S_{\omega,g}  &  n^X_{\omega,1} &  \cdots & n^X_{\omega,g} \\
						\end{array} \right],
\end{equation}

@@

\begin{equation}
	\tilde{\mbf{n}}(t) =  \left[
								\begin{array}{c}
									\mbf{n}^{S}_{AA}(t) \\
									\mbf{n}^{S}_{Aa}(t) \\
									\mbf{n}^{S}_{aa}(t) \\ \hline
									\mbf{n}^{X}_{AA}(t) \\
									\mbf{n}^{X}_{Aa}(t) \\
									\mbf{n}^{X}_{aa}(t) \\ 
						\end{array} \right].
\end{equation}

@@

\begin{equation}
	\tilde{\mbf{A}}[\tilde{\mbf{n}}] = 
			\underbrace{\left(
			\begin{array}{c|c}
				\mcal{U}^S & \mbf{0} \\ \hline
				\mbf{0} & \mcal{U}^X \\
			\end{array} \right)}_{\tilde{\mbf{U}}} + 
			\underbrace{\left(
			\begin{array}{c|c}
				\mcal{F}^S(\tilde{\mbf{p}}) (1 - \delta) & \mcal{F}^S(\tilde{\mbf{p}}) (1 - \delta) \\ \hline
				\mcal{F}^X(\tilde{\mbf{p}}) & \mcal{F}^X(\tilde{\mbf{p}}) \\
			\end{array} \right)}_{\tilde{\mbf{F}}}
\end{equation}

@@

\noindent and corresponding matrices $\mbb{U}^X$, $\mbb{F}^S$, $\mbb{F}^X$. Because individuals cannot change their genotype once they are born, the survival matrices are block diagonal; $\mcal{U}^S = \mbb{U}^S$ and $\mcal{U}^X = \mbb{U}^X$. We also construct

@@

\begin{align*}
	&\tilde{\mbf{A}}[\tilde{\mbf{n}}] = 
		 \underbrace{\left(
			\begin{array}{ccc|ccc}
				\mcal{U}^S_{AA} & \mbf{0} & \mbf{0} & \mbf{0} & \mbf{0} & \mbf{0} \\ 
				\mbf{0} & \mcal{U}^S_{Aa} & \mbf{0} & \mbf{0} & \mbf{0} & \mbf{0} \\
				\mbf{0} & \mbf{0} & \mcal{U}^S_{Aa} & \mbf{0} & \mbf{0} & \mbf{0} \\ \hline
				\mbf{0} & \mbf{0} & \mbf{0} & \mcal{U}^X_{AA} & \mbf{0} & \mbf{0} \\
				\mbf{0} & \mbf{0} & \mbf{0} & \mbf{0} & \mcal{U}^X_{aa} & \mbf{0}\\
				\mbf{0} & \mbf{0} & \mbf{0} & \mbf{0} & \mbf{0} & \mcal{U}^X_{aa}
			\end{array} \right)}_{\tilde{\mbf{U}}} + \\ 
	&\underbrace{\left(
			\begin{array}{ccc|ccc}
				C_{AA} (1 - \delta) \mbf{F}_{AA} & \frac{1}{4} C_{Aa} (1 - \delta) \mbf{F}_{Aa} & \mbf{0} & C_{AA} (1 - \delta) \mbf{F}_{AA} & \frac{1}{4} C_{Aa} (1 - \delta) \mbf{F}_{Aa} & \mbf{0} \\ 
				\mbf{0} & \frac{1}{2} C_{Aa} (1 - \delta) \mbf{F}_{Aa} & \mbf{0} & \mbf{0} & \frac{1}{2} C_{Aa} (1 - \delta) \mbf{F}_{Aa} & \mbf{0}  \\
				\mbf{0} & \frac{1}{4} C_{Aa} (1 - \delta) \mbf{F}_{Aa} & C_{aa} (1 - \delta) \mbf{F}_{aa} & \mbf{0} & \frac{1}{4} C_{Aa} (1 - \delta) \mbf{F}_{Aa} & C_{aa} (1 - \delta) \mbf{F}_{aa} \\ \hline
				q^{\prime}_{A} \mbf{F}_{AA} (1 - C_{AA})& \frac{1}{2} q^{\prime}_{A} \mbf{F}_{Aa} (1 - C_{Aa}) & \mbf{0} & q^{\prime}_{A} \mbf{F}_{AA} (1 - C_{AA})& \frac{1}{2} q^{\prime}_{A} \mbf{F}_{Aa} (1 - C_{Aa}) & \mbf{0} \\
				q^{\prime}_{a} \mbf{F}_{AA} (1 - C_{AA}) & \frac{1}{2} \mbf{F}_{Aa} (1 - C_{Aa}) & q^{\prime}_{A} \mbf{F}_{aa} (1 - C_{aa}) & q^{\prime}_{a} \mbf{F}_{AA} (1 - C_{AA}) & \frac{1}{2} \mbf{F}_{Aa} (1 - C_{Aa}) & q^{\prime}_{A} \mbf{F}_{aa} (1 - C_{aa}) \\
				\mbf{0} & \frac{1}{2} q^{\prime}_{a} \mbf{F}_{Aa} (1 - C_{Aa})& q^{\prime}_{a} \mbf{F}_{aa} (1 - C_{aa}) & \mbf{0} & \frac{1}{2} q^{\prime}_{a} \mbf{F}_{Aa} (1 - C_{Aa})& q^{\prime}_{a} \mbf{F}_{aa} (1 - C_{aa})
			\end{array} \right)}_{\tilde{\mbf{F}}} \numberthis
\end{align*}

@@

As a first pass at implementing the model, we construct and analyze a genotype $\times$stage-classified model for a hypothetical species with intralocus sexual conflict via the two sex functions. Our hypothetical species has two life stages: juveniles and adults. Suppose that allele $A$ is beneficial for females but detrimental for males, and that allele $a$ has the reverse effect. We also assume that the effects act only during the adult stage through reproductive success. The allele does not affect survival and transition rates. However, the survival matrices can be used to model 'late-acting' inbreeding depression by allowing different stage-specific survival and transition rates for individuals produced by self-fertilization vs.~outcrossing. By contrast, the parameter $\sigma$ only affects inbreeding depression through viability of selfed ovlues. With this in mind, we include different survival matrices for individuals produced by selfing and outcrossing as follows:

\begin{align*}
	\mbf{U}^S = &\left(
					\begin{array}{cc}
						\sigma^S_J(1 - \gamma^S) & 0 \\
						\sigma^S_J \gamma      & \sigma^S_A
					\end{array}
				\right) \\
	\mbf{U}^X = &\left(
					\begin{array}{cc}
						\sigma^X_J(1 - \gamma^X) & 0 \\
						\sigma^X_J \gamma      & \sigma^X_A
					\end{array}
				\right)
\end{align*}

@@

\begin{equation}
	\mbf{F}_{ii} = \left(
					\begin{array}{cc}
						0 & C_{ii}(1 - \delta) f_{ii} \\
						0 & 0
					\end{array}
				\right),
\end{equation}
\noindent and
\begin{equation}
	\mbf{F}^{\prime}_{ii} = \left(
					\begin{array}{cc}
						0 & (1 - C_{ii}) f^{\prime}_{ii} \\
						0 & 0
					\end{array}
				\right),
\end{equation}

@@

\begin{align*}
	\tilde{\mbf{A}}[\tilde{\mbf{n}}] = 
			&\underbrace{\left(
			\begin{array}{c|c}
				\mcal{U}^S & \mbf{0} \\ \hline
				\mbf{0} & \mcal{U}^X \\
			\end{array} \right)}_{\tilde{\mbf{U}}} + 
			\underbrace{\left(
			\begin{array}{c|c}
				\mcal{F}^S(\tilde{\mbf{p}}) C(1 - \delta) & \mcal{F}^S(\tilde{\mbf{p}}) C(1 - \delta) \\ \hline
				\mcal{F}^X(\tilde{\mbf{p}}) (1 - C) & \mcal{F}^X(\tilde{\mbf{p}}) (1 - C)\\
			\end{array} \right)}_{\tilde{\mbf{F}}} \\
			= & \left(\begin{array}{c|c}
				\mcal{U}^S + \mcal{F}^S(\tilde{\mbf{p}}) C(1 - \delta) & \mcal{F}^S(\tilde{\mbf{p}}) C(1 - \delta) \\ \hline
				\mcal{F}^X(\tilde{\mbf{p}})(1 - C) & \mcal{U}^X + \mcal{F}^X(\tilde{\mbf{p}}) (1 - C)
			\end{array} \right) \numberthis
\end{align*}

\begin{table}[htbp]
 \centering
 \caption{\bf Fitness expressions for Sexually Antagonistic selection in hermaphrodites}
\begin{tabular}{lccc}
 \toprule
					&  \multicolumn{3}{c}{{\textit{Genotype}}} \\ 
\cline{2-4}
					& $AA$			& $Aa$ 					& $aa$ 		\\ \hline
Female function:	& $f_{AA}$		& $f_{Aa} - h_f s_f$	& $f_{aa} - s_f$ \\	
Male function:		& $f_{AA} - s_m$& $f_{Aa} - h_m s_m$	& $f_{aa}$ 		\\	
\hline
\end{tabular}
\end{table}
